Police Shootings and Racism in America

### 2020-10-19 police_eda.Rmd
added last of eda i hope

```{r}
temp <- final_df %>% 
  summarise(Black = sum(str_count(OfficerRace, "B")), 
            White = sum(str_count(OfficerRace, "W")),
            Hispanic = sum(str_count(OfficerRace, "L")),
            Asian = sum(str_count(OfficerRace, "A")), 
            Other = sum(str_count(OfficerRace, "O")))
ggplot(temp, aes(x = ))
```

```{r}
temp <- final_df %>% 
  filter(sb == 1) %>% 
  group_by(fatal) %>% 
  summarise(ob = sum(ob == 1),
            ow = sum(ow == 1),
            ol = sum(ol == 1),
            oa = sum(oa == 1),
            oo = sum(oo == 1)) %>% 
  mutate(ob = ob/(ob + ow + ol + oa + oo), 
         ow = ow/(ob + ow + ol + oa + oo),
         ol = ol/(ol + ow + ol + oa + oo), 
         oa = oa/(ob + ow + ol + oa + oo), 
         oo = oo/(ob + ow + ol + oa + oo))
temp <- rename(temp, Black = ob, White = ow, Latino = ol, Asian = oa, Other = oo)

@@

officerplot <- ggplot(temp2, aes(x=race, y = value, fill = variable)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions for Fatal vs. Non-Fatal Police Shootings Where Black Victims were Present") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

@@

>>>>

```{r}
temp <- final_df %>% 
  summarise(Black = sum(str_count(OfficerRace, "B")), 
            White = sum(str_count(OfficerRace, "W")),
            Hispanic = sum(str_count(OfficerRace, "L")),
            Asian = sum(str_count(OfficerRace, "A")), 
            Other = sum(str_count(OfficerRace, "O")))

temp2 <- temp %>%
    t() %>%
    as.data.frame(stringsAsFactors = F) %>%
    rownames_to_column("race") %>%
    `rownames<-`(NULL)
temp2 <- rename(temp2, Count = 2) %>% 
  mutate(total = Count/sum(Count))

officerplot2 <- ggplot(temp2, aes(x=race, y = total)) +
  geom_bar(position = "dodge", stat="identity", fill ="#D48C7D") + 
  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions of Officers for All Shootings") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size=10, hjust = 0.5))
ggsave(filename = "figures/officerprop.png", officerplot2)
```

```{r sb-officer}
temp <- final_df %>% 
  filter(sb == 1) %>% 
  group_by(fatal) %>% 
  summarise(ob = sum(ob == 1),
            ow = sum(ow == 1),
            ol = sum(ol == 1),
            oa = sum(oa == 1),
            oo = sum(oo == 1)) %>% 
  mutate(ob = ob/(ob + ow + ol + oa + oo), 
         ow = ow/(ob + ow + ol + oa + oo),
         ol = ol/(ol + ow + ol + oa + oo), 
         oa = oa/(ob + ow + ol + oa + oo), 
         oo = oo/(ob + ow + ol + oa + oo))
temp <- rename(temp, Black = ob, White = ow, Latino = ol, Asian = oa, Other = oo)

@@

officerplot <- ggplot(temp2, aes(x=race, y = value, fill = variable)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions of Officers Present at Fatal vs. Non-Fatal Police Shootings Involving Black Victims") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

@@

Nothing too crazy happening here.

### 2020-10-19 submissionii.Rmd
added last of eda i hope

|       Plotted below in \emph{Figure 1} are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs. the race proportions of those cities' total populations in 2013. Because the bars representing shootings in the Black race category are taller than the bars representing city-wide population in all 4 plots, it is evident Black people are disproportionately the victims of police shootings compared to city race proportions. This can especially be seen in Houston, where Black people make up less than 25% of the city population yet more than 50% of the victim population of police shootings.
```{r, figures-city, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
knitr::include_graphics("figures/cityeda1.png")
knitr::include_graphics("figures/cityeda2.png")
```

```{r num_rows, eval = F}
final_eda <- readRDS('data/final_eda.rds')

@@

mult_victims <- final_eda %>% 
  filter(NumberOfSubjects > 1)
nrow(mult_victims)
```

```{r, figures-single, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
par(mar = c(2, 2, .1, .1))
knitr::include_graphics("figures/singleeda1.png")
knitr::include_graphics("figures/singleeda2.png")
```
|       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among all shootings with only one victim, it appears in \emph{Figure 2} that the victim is more often Black than some other race. However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally. Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small. This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal. 
|       Similarly, Black victims make up the highest proportion of victims in shootings with multiple victims. However, in these shootings, all race types are more likely to be shot non-fatally than fatally, with Latinos having the highest proportion of being shot fatally out of all other race types. Additionally, it should be noted that there are no victims with race type "Other" in shootings where there are multiple victims. These graphs can be found as \emph{Figure 3} in the Appendix.

@@

```{r, figures-btw, fig.show="hold", out.width="50%", fig.cap="City Black to White Ratios Distributions for Fatal vs. Non-Fatal Police Shootings"}
par(mar = c(2, 2, .1, .1))
knitr::include_graphics("figures/btweda.png")
```

Means are not that different, will probably just use percent black and percent white then.

>>>>

```{r, figures-city, fig.show="hold", out.width="25%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
knitr::include_graphics("figures/cityeda1.png")
knitr::include_graphics("figures/cityeda2.png")
knitr::include_graphics("figures/cityeda3.png")
knitr::include_graphics("figures/cityeda4.png")
```
|       Plotted above in \emph{Figure 1} are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs. the race proportions of those cities' total populations in 2013. Because the bars representing shootings in the Black race category are taller than the bars representing city-wide population in all 4 plots, it is evident Black people are disproportionately the victims of police shootings compared to city race proportions. This can especially be seen in Houston, where Black people make up less than 25% of the city population yet more than 50% of the victim population of police shootings.
```{r num_rows, eval = F}
final_eda <- readRDS('data/final_eda.rds')

@@

mult_victims <- final_eda %>% 
  filter(NumberOfSubjects > 1)
nrow(mult_victims)
```
```{r, figures-single, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
par(mar = c(2, 2, .1, .1))
knitr::include_graphics("figures/singleeda1.png")
knitr::include_graphics("figures/singleeda2.png")
```
|       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among all shootings with only one victim, it appears in \emph{Figure 2} that the victim is more often Black than some other race. However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally. Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small. This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal. 
|       Similarly, Black victims make up the highest proportion of victims in shootings with multiple victims. However, in these shootings, all race types are more likely to be shot non-fatally than fatally, with Latinos having the highest proportion of being shot fatally out of all other race types. Additionally, it should be noted that there are no victims with race type "Other" in shootings where there are multiple victims. These graphs can be found as \emph{Figure 3} in the Appendix.

@@

```{r, figures-btw, fig.show="hold", out.width="50%", fig.cap="Proportion of Fatal Shootings vs. Black to White Ratios in Cities in Dataset (left) and vs. Black Population Percentages in Cities in Dataset (right)"}
par(mar = c(2, 2, .1, .1))
knitr::include_graphics("figures/btweda.png")
knitr::include_graphics("figures/bpropeda.png")
```

```{r, figures-officersbprop, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Officers For All Shootings (left) and For Fatal vs. Non-Fatal Police Shootings Where At Least One Black Victim was Present (right) and "}
par(mar = c(2, 2, .1, .1))
knitr::include_graphics("figures/officerprop.png")
knitr::include_graphics("figures/officerpropsb.png")
```
